---
title: "Assignment_3"
output: word_document
date: "2023-10-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Install and import libraries
```{r}
# Load necessary libraries
options(warn = -1)  # Suppress all warnings
if (!require("caret")) install.packages("caret")
if (!require("e1071")) install.packages("e1071")
if (!require("reshape2")) install.packages("reshape2")
library(caret)
library(e1071)
library(reshape2)
```

### Load the dataset
```{r}
# Load the data
data <-  read.csv("~/Downloads/UniversalBank.csv")
```

### partition of data into training and validation sets

```{r}
set.seed(123)
index <- createDataPartition(data$Personal.Loan, p = 0.6, list = FALSE)
train_data <- data[index, ]
val_data <- data[-index, ]
colnames(train_data)

```

#A: Create a pivot table for the training data with Online as a column variable, Credit Card as a row variable, and Personal Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table().
```{r}
# Melt the data, specifying 'variable.name' and 'value.name'
data_melted <- melt(train_data, id.vars = c("Online", "CreditCard"), measure.vars = "Personal.Loan", variable.name = "Variable", value.name = "Personal_Loan")

# Cast the melted data into a wider format for the pivot table
pivot <- dcast(data_melted, CreditCard + Personal_Loan ~ Online, fun.aggregate = length)

print(pivot)


```

#B: Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Personal Loan = 1) conditional on having a bank credit card (CreditCard = 1) and being an active user of online banking services (Online = 1)].

```{r}

# Probability calculation for part b
subset_data <- subset(train_data, CreditCard == 1 & Online == 1)
prob_b <- nrow(subset(subset_data, `Personal.Loan` == 1)) / nrow(subset_data)

print(prob_b)
```

#C: Create two separate pivot tables for the training data. One will have Personal Loan (rows) as a function of Online (columns) and the other will have Personal Loan (rows) as a function of CreditCard.
```{r}
# Pivot tables for part c
pivot_loan_online <- table(train_data$`Personal.Loan`, train_data$Online)
pivot_loan_cc <- table(train_data$`Personal.Loan`, train_data$CreditCard)

print(pivot_loan_online)
print(pivot_loan_cc)

```

#D: Compute the following quantities [P(A | B) means “the probability of A given B”]:
```{r}
# Probabilities for part d
P_CC_given_loan_1 <- pivot_loan_cc[2,2] / sum(pivot_loan_cc[2,])
P_online_given_loan_1 <- pivot_loan_online[2,2] / sum(pivot_loan_online[2,])
P_loan_1 <- sum(pivot_loan_cc[2,]) / sum(pivot_loan_cc)
P_CC_given_loan_0 <- pivot_loan_cc[1,2] / sum(pivot_loan_cc[1,])
P_online_given_loan_0 <- pivot_loan_online[1,2] / sum(pivot_loan_online[1,])
P_loan_0 <- sum(pivot_loan_cc[1,]) / sum(pivot_loan_cc)

```

#E: Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1) 
```{r}

# Naive Bayes probability for part e
total <- nrow(train_data)
P_CC_1 <- sum(train_data$CreditCard == 1) / total
P_online_1 <- sum(train_data$Online == 1) / total
P_loan_1_given_CC_1_and_online_1 <- (P_CC_given_loan_1 * P_online_given_loan_1 * P_loan_1) / (P_CC_1 * P_online_1)

```

#F: Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?

```{r}
# 9.05 vs 2.57

# the vaule obtained from the pivot table is less accurate
```

#G: Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

```{r}
# --- Task: Train a naive Bayes model and make a prediction for P(Loan = 1 | CC = 1, Online = 1) ---
model <- naiveBayes(as.factor(`Personal.Loan`) ~ Online + CreditCard, data = train_data)
newdata <- data.frame(Online = 1, CreditCard = 1)
pred <- predict(model, newdata, type = "raw")
prob_from_model <- pred[1, "1"]  # "1" is the factor level for loan acceptance
print(prob_from_model)  # This prints the probability from the naive Bayes model

```

