---
title: "Assignment 4"
output: word_document
date: "2023-10-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(factoextra)
pharmaceutical_0data<-read.csv("~/Downloads/Pharmaceuticals.csv")
pharmaceutical_data<-na.omit(pharmaceutical_0data)

```

Using the numerical variables (1 to 9) to cluster the 21 firms.
```{r}
row.names(pharmaceutical_0data)<-pharmaceutical_0data[,1]
Clustering_data<-pharmaceutical_0data[,3:11]
```
Scaling the data according to requirement
```{r}
set.seed(143)
ScaledData<-scale(Clustering_data)

```


```{r}
k_wss<-fviz_nbclust(ScaledData,kmeans,method="wss")
k_silhouette<-fviz_nbclust(ScaledData,kmeans,method="silhouette")
k_wss
k_silhouette
distance<-dist(ScaledData,metho='euclidean')
fviz_dist(distance)
```

We will be choose silhouette k value of 5 because this will ensure that the sum of squares is low along with proper separation within the clusters.

```{r}
set.seed(143)
kmeans_5<-kmeans(ScaledData,centers = 5, nstart = 10)
kmeans_5
plot_kmeans_5<-fviz_cluster(kmeans_5,data = ScaledData) + ggtitle("K=5")
plot_kmeans_5

Clustering_data_1<-Clustering_data%>%
  mutate(Cluster_no=kmeans_5$cluster)%>%
  group_by(Cluster_no)%>%summarise_all('mean')
Clustering_data_1

```
Companies are grouped into following clusters:
Cluster_1= ABT,AHM,AZN,BMY,LLY,NVS,SGP,WYE
Cluster_2= BAY,CHTT,IVX
Cluster_3=AVE,ELN,MRX,WPI
Cluster_4=AGN,PHA
Cluster_5=GSK,JNJ,MRK,PFE
From the above clusters 
1.Cluster_1 Companies with moderate returns on equity and investment
2.Cluster_2 Companies with Poor returns on Assets(ROA), Return on Equity (ROE), Low market Capitalization, and weak Asset turnover which implies that these companies are Highly Risky
3.Cluster_3 Companies Similar to those in cluster 2 but with slightly lower levels of risk
4.Cluster_4 Companies with very high price to earnings (P/E) ratios but poor ROA and ROE, making them more risky than those in cluster 2
5.Cluster_5 Companies with high market capitalization, ROE and ROA

```{r}
Clustering_data_2<- pharmaceutical_data[,12:14] %>% mutate(Clusters=kmeans_5$cluster)
ggplot(Clustering_data_2, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')
ggplot(Clustering_data_2, mapping = aes(factor(Clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
ggplot(Clustering_data_2, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')


```

By examining the data, a distinct pattern becomes apparent concerning the connection between the clusters and the 'Median Recommendation' variable. Cluster 2 predominantly indicates recommendations spanning from 'hold' to 'moderate buy,' while Cluster 3 tends to favor recommendations from 'moderate buy' to 'moderate sell.'Upon closer inspection of the geographical distribution, it becomes evident that a substantial percentage of pharmaceutical companies are headquartered in the United States, and there is no conspicuous spatial arrangement. Nevertheless, there is no apparent association between clusters and the stock exchange, aside from the observation that the majority of these companies are publicly listed on the New York Stock Exchange (NYSE).


Naming clusters:


**Cluster 1 = Investment Cluster.**

**Cluster 2 = Tight Cluster.**

**Cluster 3 = Low Profit Cluster.**

**Cluster 4 = Risky Cluster.**

**Cluster 5 = Large Purchase Cluster.**